---
title: "Comparison of DNA isolation methods, Library preparation protocols, Bionformatic workflows and Databases on Toti samples"
author: Balázs Kakuk
output:
   html_document:
      toc: TRUE
      toc_depth: 3
      toc_float: TRUE
      number_sections: TRUE
      df_print: kable
      
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.align = 'center', fig.height = 10, fig.width = 30, warning = F, message = F, error = F)
library(tidyverse)
library(ggpubr)
library(factoextra)
library(FactoMineR)
library(ggrepel)
library(phyloseq, quietly = T)
library(DESeq2)
library(data.table)
library(ggsci)
library(hrbrthemes)
library(ggh4x)
library(RColorBrewer)
library(stringr); library(stringi)

### Own functions
for(f in list.files('E:/my.R.packages/Rlyeh-main/R', full.names = T)) {source(f)}
for(f in list.files('E:/my.R.packages/minitax/R', full.names = T)) {source(f)}

rownames.from.1st.col <- function(x) {
  x <- data.frame(x[,-1], row.names = x[,1])
  return(x)
}


### General settings

filt.version <- ''
filtering    <- ''
writetables  <- T
save.Figs    <- F
res.dir      <- 'Toti_all_comp_minitax_NCBI';dir.create(res.dir)
#platforms <- c('illumina', 'PacBIO')

## color palette
alpha   <- 0.9
pal.man <- c(brewer.pal(7, 'Set1'), brewer.pal(8, 'Accent')[c(1,2,6)], brewer.pal(8, 'Set2')[c(1,3,5,6,7)], brewer.pal(8, 'Dark2') )
pal.man <- c(colorvec[c(8,32,12:14,17,24,1:2,4,7,5,6,3,10,11,27:28,9,29)], 'grey', 'black')
pal.man <- alpha(c(pal.man, pal.man), alpha)

## taxonomic ranks
ranks <- c("superkingdom", "phylum", "class", "order", "family", "genus", "species")

### minitax settings
## Databse
db <- 'all_NCBI_genomes'
db.dir <- 'E:/data/databases/all_NCBI_genomes'
## Genome Size normalization?
NormToGenomeSize <- F
NormFunction     <- 'mean'
## MAPQ filtering ?
mapq.filt <- NA #  1:59 # 
## Filtering samples
filter.samples.pattern <- 'LCA|Undetermined|SpeciesEstimate'  ### filter out LCA method


## Exclude unclassified reads?
exclude.unclass <- T

## Omit very low abundance taxa
abund.threshold <- 10

### Gold Standard Settings
cami.lineage <- 'db'
add.GS.taxa <- F ## add all GS taxa to the barplots and the correlation plots?
zymo.GS <- read_delim('E:/data/databases/zymo/zymo_6331_composition.txt')
GS_name <- 'GS_reference'
GS_nr   <- 1
just_GS <- T
vars <- 'platform' #('project') # c('DNA_isolation_method', 'db', 'workflow', 'method', 
glom_spec <- T
consider.euk <- T
count_multiplier <- 100000
count_colname    <- 'Genome Copy'

## Fix taxa
crop_taxa <- NULL # 'Veillonella rogosae|Prevotella corporis'

genera_to_glom <- NULL # data.frame(
#  glom= c('Veillonella', 'Faecalibacterium', 'Prevotella'),
#  to  = c('Veillonella rogosae','Faecalibacterium prausnitzii', 'Prevotella corporis')) ## NULL

species_to_glom <- NULL # data.frame(
#  glom= c('Bacteroides fragilis', 'Bacteroides fragilis CAG:47', 'Bacteroides fragilis_A', "Secundilactobacillus pentosiphilus", "Clostridium hathewayi CAG:224"),
#  to  = c(rep('Bacteroides fragilis', 3), 'Lactobacillus pentosiphilus', 'Hungatella hathewayi' )) ## NULL



```



# Notes

- Minitax was run in BestAln method
- In minitax, reads were *NOT* filtered for MAPQ
- In minitax, results were *NOT* normalized to Genome Size
- Unclassified reads *were* excluded
- Taxa with total reads number of 10 or lower were excluded 
- Eukaryotes were *NOT* excluded from the analyis
- Same samples for different sequencing dates were combined


```{r, import}

ps.all <- readRDS('Toti_all_PS.summed.filtered.rds')

metadata <- data.frame(sample_data(ps.all))
otutab   <- data.frame(otu_table(ps.all))
taxtab   <- data.frame(tax_table(ps.all))

```





#  Relative abundance of taxa at each taxonomic level

```{r, Barplots}

## Filter for samples
samples.to.keep <- plyr::rbind.fill(
  
  metadata[metadata$workflow == 'minitax'  & 
           metadata$db       == 'all_NCBI_genomes'
           #metadata$DNA_isolation_method == 'I'
           , ]
             
)$sample

ps.all <- prune_samples(samples.to.keep, ps.all) # ps.all # 


## Filter for taxa below threshold
if (!is.na(abund.threshold)) {

  otutab   <- data.frame(otu_table(ps.all))
  taxtab   <- data.frame(tax_table(ps.all))
  metadata <- data.frame(sample_data(ps.all))

  otutab   <- otutab %>% dplyr::filter(rowSums(select(., everything())) >= abund.threshold)
  taxtab   <- taxtab[is.element(rownames(taxtab), rownames(otutab)), ]

  ps.all   <- phyloseq(otu_table(as.matrix(otutab), taxa_are_rows = T),
                       tax_table(as.matrix(taxtab)),
                       sample_data(metadata))


}

ps <- ps.all

scales <- 'fixed'
fig.height  <- 9
fig.width   <- 14
top <- 20
top.MCM <- 8

multiV <- c(F, T)

fig.height.vr <- fig.height
fig.width.vr  <- fig.width

## plotting function for barplots
plotfun <- function(ps, x='sample_nr', t=t, fill=t) {
  plot_bar(ps, x=x, fill=fill
         ) + 
    scale_fill_manual(values=as.character(pal.man)) +
    guides(fill = guide_legend(direction = "vertical", ncol = 1)) +
    theme_ipsum() + 
    theme(legend.position="right", 
          panel.spacing.x = unit(1.5, 'mm'),
          axis.text.x = element_text(),
          axis.title.x = element_blank()) +
    # facet_grid(rows = vars(sequencing_date), cols = vars(DNA_isolation_method), scales = 'free')
    facet_nested(cols   = vars(Vregion, platform, db, workflow),
                 rows   = vars(DNA_isolation_method),
                 scales = 'free_x')
}

```


## Phylum level

```{r, Glom taxa at Phylum-level}
n <- 2

t <- rank_names(ps)[n]

## glomerate at n taxonomic level, regardless of lineage
ps.glom <- glom_taxa_n(ps, n=n, ignore_lineage = T)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## get data
ps.data.glom.top  <- df.from.ps(ps.glom.top,   comb=F)
ps.data.glom      <- df.from.ps(ps.glom,       comb=F)

## filenames for plots and tables
filenames <- c(
  paste0(res.dir, '_top' , top, '_', t, '_ratios'),
  paste0(res.dir, '_all_', t,           '_ratios') 
)
## write tables
if (writetables){
 write_tsv(ps.data.glom.top, paste0(res.dir, '/', filenames[1], '.tsv')) 
 write_tsv(ps.data.glom,     paste0(res.dir, '/', filenames[2], '.tsv')) 
}

```


```{r, Phylum-level plot, fig.height=fig.height.vr, fig.width = fig.width.vr, eval=multiV[2]}

ggtop20    <- plotfun(
  ps.glom.norm.top,
                t=t, fill=t) + 
  theme(legend.position = 'right') +
  ggtitle(paste0('Top', top, ' ', t))

ggtop20

if (save.Figs){
  ggsave(paste0(res.dir, '/', filenames[1], '_Vregions.jpg'), width = fig.width.vr, height = fig.height.vr)
}



ps.glom.norm.top.phy <- ps.glom.norm.top
ps.glom.norm.phy     <- ps.glom.norm
ps.glom.top.phy      <- ps.glom.top
ps.glom.phy          <- ps.glom


```


## Order level

```{r, Glom taxa at Order-level}
n <- 4

t <- rank_names(ps)[n]

## glomerate at n taxonomic level, regardless of lineage
ps.glom <- glom_taxa_n(ps, n=n, ignore_lineage = T)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
 top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## get data
ps.data.glom.top  <- df.from.ps(ps.glom.top,   comb=F)
ps.data.glom      <- df.from.ps(ps.glom,       comb=F)

## filenames for plots and tables
filenames <- c(
  paste0(res.dir, '_top' , top, '_', t, '_ratios'),
  paste0(res.dir, '_all_', t,           '_ratios') 
)
## write tables
if (writetables){
 write_tsv(ps.data.glom.top, paste0(res.dir, '/', filenames[1], '.tsv')) 
 write_tsv(ps.data.glom,     paste0(res.dir, '/', filenames[2], '.tsv')) 
}

```

```{r, Order-level plot, fig.height = fig.height.vr, fig.width = fig.width.vr, eval=multiV[2]}


ggtop20    <- plotfun(
  ps.glom.norm.top,
                t=t, fill=t) + 
  theme(legend.position = 'right') +
  ggtitle(paste0('Top', top, ' ', t))

ggtop20

if (save.Figs){
  ggsave(paste0(res.dir, '/', filenames[1], '_Vregions.jpg'), width = fig.width.vr, height = fig.height.vr)
}



ps.glom.norm.top.ord <- ps.glom.norm.top
ps.glom.norm.ord     <- ps.glom.norm
ps.glom.top.ord      <- ps.glom.top
ps.glom.ord          <- ps.glom

```

## Family level

```{r, Glom taxa at Family-level}
n <- 5

t <- rank_names(ps)[n]

## glomerate at n taxonomic level, regardless of lineage
ps.glom <- glom_taxa_n(ps, n=n, ignore_lineage = T)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## get data
ps.data.glom.top  <- df.from.ps(ps.glom.top,   comb=F)
ps.data.glom      <- df.from.ps(ps.glom,       comb=F)

## filenames for plots and tables
filenames <- c(
  paste0(res.dir, '_top' , top, '_', t, '_ratios'),
  paste0(res.dir, '_all_', t,           '_ratios') 
)
## write tables
if (writetables){
 write_tsv(ps.data.glom.top, paste0(res.dir, '/', filenames[1], '.tsv')) 
 write_tsv(ps.data.glom,     paste0(res.dir, '/', filenames[2], '.tsv')) 
}

```

```{r, Family-level plot, fig.height = fig.height.vr, fig.width = fig.width.vr, eval=multiV[2]}

ggtop20    <- plotfun(
  ps.glom.norm.top,
                t=t, fill=t) + 
  theme(legend.position = 'right') +
  ggtitle(paste0('Top', top, ' ', t))

ggtop20

if (save.Figs){
  ggsave(paste0(res.dir, '/', filenames[1], '_Vregions.jpg'), width = fig.width.vr, height = fig.height.vr)
}



ps.glom.norm.top.fam <- ps.glom.norm.top
ps.glom.norm.fam     <- ps.glom.norm
ps.glom.top.fam      <- ps.glom.top
ps.glom.fam          <- ps.glom

```


## Genus level

```{r, Glom taxa at Genus-level}
n <- 6

t <- rank_names(ps)[n]

## glomerate at n taxonomic level, regardless of lineage
ps.glom <- glom_taxa_n(ps, n=n, ignore_lineage = T)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## get data
ps.data.glom.top  <- df.from.ps(ps.glom.top,   comb=F)
ps.data.glom      <- df.from.ps(ps.glom,       comb=F)

## filenames for plots and tables
filenames <- c(
  paste0(res.dir, '_top' , top, '_', t, '_ratios'),
  paste0(res.dir, '_all_', t,           '_ratios') 
)
## write tables
if (writetables){
 write_tsv(ps.data.glom.top, paste0(res.dir, '/', filenames[1], '.tsv')) 
 write_tsv(ps.data.glom,     paste0(res.dir, '/', filenames[2], '.tsv')) 
}

```

```{r, Genus-level plot, fig.height = fig.height.vr, fig.width = fig.width.vr, eval=multiV[2]}

ggtop20    <- plotfun(
  ps.glom.norm.top,
                t=t, fill=t) + 
  theme(legend.position = 'right') +
  ggtitle(paste0('Top', top, ' ', t))

ggtop20

if (save.Figs){
  ggsave(paste0(res.dir, '/', filenames[1], '_Vregions.jpg'), width = fig.width.vr, height = fig.height.vr)
}



ps.glom.norm.top.gen <- ps.glom.norm.top
ps.glom.norm.gen     <- ps.glom.norm
ps.glom.top.gen      <- ps.glom.top
ps.glom.gen          <- ps.glom

```


## Species level

```{r, Glom taxa at Species-level}
n <- 7

t <- rank_names(ps)[n]

## glomerate at n taxonomic level, regardless of lineage
#if (!all(taxa_names(ps) == data.frame(ps@tax_table$species))) {
  ps.glom <- glom_taxa_n(ps, n=n, ignore_lineage = T)  
#} else {
#  ps.glom <- ps
#}


## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## normalize PS object per taxon sums
ps.glom.norm       <- norm.ps(ps.glom, 'sum')

## top20 taxa from the glomerated at specific tax level
top.taxa           <- names(sort(taxa_sums(ps.glom.norm), decreasing=TRUE))[1:top]

## add all Gold standard taxa?
if (add.GS.taxa) {
  top.taxa  <- unique(c(top.taxa, taxtab.gs[,t]))
}

## prune normalized taxa
ps.glom.norm.top   <- prune_taxa(top.taxa, ps.glom.norm)
## prune taxa
ps.glom.top        <- prune_taxa(top.taxa, ps.glom)

## get data
ps.data.glom.top  <- df.from.ps(ps.glom.top,   comb=F)
ps.data.glom      <- df.from.ps(ps.glom,       comb=F)

## filenames for plots and tables
filenames <- c(
  paste0(res.dir, '_top' , top, '_', t, '_ratios'),
  paste0(res.dir, '_all_', t,           '_ratios') 
)
## write tables
if (writetables){
 write_tsv(ps.data.glom.top, paste0(res.dir, '/', filenames[1], '.tsv')) 
 write_tsv(ps.data.glom,     paste0(res.dir, '/', filenames[2], '.tsv')) 
}

```

```{r, Species-level plot, fig.height = fig.height.vr, fig.width = fig.width.vr, eval=multiV[2]}

ggtop20    <- plotfun(
  ps.glom.norm.top,
                t=t, fill=t) + 
  theme(legend.position = 'right') +
  ggtitle(paste0('Top', top, ' ', t))

ggtop20

if (save.Figs){
  ggsave(paste0(res.dir, '/', filenames[1], '_Vregions.jpg'), width = fig.width.vr, height = fig.height.vr)
}



ps.glom.norm.top.spec <- ps.glom.norm.top
ps.glom.norm.spec     <- ps.glom.norm
ps.glom.top.spec      <- ps.glom.top
ps.glom.spec          <- ps.glom

```



```{r}
save.image(paste0(res.dir, '.RData'))
```



```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```

