---
title: "Statistical comparison of DNA Isolation Methods across each Library preparation protocol, Toti samples - analyzed with minitax on all NCBI genomes DB"
author: Balázs Kakuk
output:
   html_document:
      toc: TRUE
      toc_depth: 3
      toc_float: TRUE
      number_sections: TRUE
      df_print: kable
      
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.align = 'center', fig.height = 6, fig.width = 9, warning = F, message = F, error = F)
library(tidyverse)
library(ggpubr)
library(factoextra)
library(FactoMineR)
library(ggrepel)
library(phyloseq, quietly = T)
library(DESeq2)
library(data.table)
library(ggsci)
library(hrbrthemes)
library(ggh4x)
library(RColorBrewer)
library(stringr); library(stringi)
library(vegan)

### Load data
load('Toti_all_comp.RData')


### Own functions
for(f in list.files('E:/my.R.packages/Rlyeh-main/R', full.names = T)) {source(f)}
for(f in list.files('E:/my.R.packages/minitax/R', full.names = T)) {source(f)}

rownames.from.1st.col <- function(x) {
  x <- data.frame(x[,-1], row.names = x[,1])
  return(x)
}


### General settings

filt.version <- ''
filtering    <- ''
writetables  <- T
save.Figs    <- F
res.dir      <- 'Toti_all_comp_DNA.Iso.Methods_statistics';dir.create(res.dir)
#platforms <- c('illumina', 'PacBIO')

## color palette
alpha   <- 0.9
pal.man <- c(brewer.pal(7, 'Set1'), brewer.pal(8, 'Accent')[c(1,2,6)], brewer.pal(8, 'Set2')[c(1,3,5,6,7)], brewer.pal(8, 'Dark2') )
pal.man <- c(colorvec[c(8,32,12:14,17,24,1:2,4,7,5,6,3,10,11,27:28,9,29)], 'grey', 'black')
pal.man <- alpha(c(pal.man, pal.man), alpha)

## taxonomic ranks
ranks <- c("superkingdom", "phylum", "class", "order", "family", "genus", "species")

### minitax settings
## Databse
db <- 'all_NCBI_genomes'
db.dir <- 'E:/data/databases/all_NCBI_genomes'
## Genome Size normalization?
NormToGenomeSize <- F
NormFunction     <- 'mean'
## MAPQ filtering ?
mapq.filt <- NA #  1:59 # 
## Filtering samples
filter.samples.pattern <- 'LCA|Undetermined|SpeciesEstimate'  ### filter out LCA method


## Exclude unclassified reads?
exclude.unclass <- T

## Omit very low abundance taxa
abund.threshold <- 10

### Gold Standard Settings
cami.lineage <- 'db'
add.GS.taxa <- F ## add all GS taxa to the barplots and the correlation plots?
zymo.GS <- read_delim('E:/data/databases/zymo/zymo_6331_composition.txt')
GS_name <- 'GS_reference'
GS_nr   <- 1
just_GS <- T
vars <- 'platform' #('project') # c('DNA_isolation_method', 'db', 'workflow', 'method', 
glom_spec <- T
consider.euk <- T
count_multiplier <- 100000
count_colname    <- 'Genome Copy'

## Fix taxa
crop_taxa <- NULL # 'Veillonella rogosae|Prevotella corporis'

genera_to_glom <- NULL # data.frame(
#  glom= c('Veillonella', 'Faecalibacterium', 'Prevotella'),
#  to  = c('Veillonella rogosae','Faecalibacterium prausnitzii', 'Prevotella corporis')) ## NULL

species_to_glom <- NULL # data.frame(
#  glom= c('Bacteroides fragilis', 'Bacteroides fragilis CAG:47', 'Bacteroides fragilis_A', "Secundilactobacillus pentosiphilus", "Clostridium hathewayi CAG:224"),
#  to  = c(rep('Bacteroides fragilis', 3), 'Lactobacillus pentosiphilus', 'Hungatella hathewayi' )) ## NULL



```



# Notes

- Minitax was run in BestAln method
- In minitax, reads were *NOT* filtered for MAPQ
- In minitax, results were *NOT* normalized to Genome Size
- Unclassified reads *were* excluded
- Taxa with total reads number of 10 or lower were excluded 
- Eukaryotes were *NOT* excluded from the analyis
- Same samples for different sequencing dates were combined
- The statistical tests were carried out on sample-wise Bray-Curtis distances


```{r, prepare}

metadata <- data.frame(sample_data(ps.all))

samples.to.keep <- plyr::rbind.fill(
  metadata[metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample


ps.all <- prune_samples(samples.to.keep, ps.glom.norm) # use the 100% normalized data
# OR:
# ps.all <- prune_samples(samples.to.keep, ps.all) # use the original count data -> worse results


### Filter for outliers
rm.outliers <- F
if (rm.outliers) {
ps.all <- prune_samples(
  sample_names(ps.all)[!is.element(sample_names(ps.all), 
  c(
    'Toti_Illumina_V1_V3_minitax_BestAln_all_NCBI_genomes_MN_4',
    'Toti_ONT_V1_V9_minitax_BestAln_all_NCBI_genomes_Q_1',
    'Toti_Illumina_V3_V4_minitax_BestAln_all_NCBI_genomes_MN_1',
    'Toti_Illumina_V3_V4_minitax_BestAln_all_NCBI_genomes_Z_4',
    'Toti_Illumina_WGS_minitax_BestAln_all_NCBI_genomes_Z_1'
  ))],
  ps.all)
}


metadata <- data.frame(sample_data(ps.all))
otutab   <- data.frame(otu_table(ps.all))
taxtab   <- data.frame(tax_table(ps.all))


write_tsv(metadata, paste0(res.dir, '/stat.test.metadata.tsv'))
write_tsv(otutab, paste0(res.dir, '/stat.test.otu_table.tsv'))
write_tsv(taxtab, paste0(res.dir, '/stat.test.tax_table.tsv'))


vregions <- c("V1_V2", "V1_V3", "V3_V4", "V1_V9", "WGS") #   unique(metadata$Vregion)
distance <- "bray"
```


# Bray-Curtis distance-based statistical tests on DNA isolation kits in each Vregion (and platform)  {.tabset}

*PERMIDISP is a multivariate analysis of the homogeneity of groups dispersions (variances), which can be used as a means of assessing beta diversity.*\

*PERMANOVA: Permutational Multivariate Analysis of Variance Using Distance Matrices*\

## V1-V2
```{r, fig.width=12, fig.height=6}
i <- 1

samples.to.keep <- plyr::rbind.fill(
  metadata[metadata$Vregion  == vregions[i] & 
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))

stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')

```

### PERMANOVA
```{r, fig.width=12, fig.height=6}
permanova <- adonis2(otutab_t ~ DNA_isolation_method, data=sampdat, method=distance)
permanova
permanova.v1v2 <- data.frame(permanova, Vregion=vregions[i])

# Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
print(pairwise_perm)
pairwise_perm.v1v2 <- data.frame(pairwise_perm, Vregion=vregions[i])

```

### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod <- betadisper(dis, groups)
mod
permidisp.v1v2    <- data.frame(distance_to_centroid=mod[["distances"]],  
                                Vregion=vregions[i])
disp.vectors.v1v2 <- data.frame(distance_to_centroid=mod[["vectors"]],
                                Vregion=paste0(vregions[i], ''))

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist.v1v2 <- data.frame(anova(mod), Vregion=vregions[i])
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest.v1v2 <- permutest(mod, pairwise = TRUE, permutations = 9999)
permutest.v1v2
permutest.v1v2 <- data.frame(permutest.v1v2[["tab"]], Vregion=vregions[i])
```

### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD <- TukeyHSD(mod))
tukey.disp.v1v2 <- data.frame(mod.HSD[["group"]], Vregion=vregions[i])
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```


## V1-V3
```{r, fig.width=12, fig.height=6}
i <- 2

samples.to.keep <- plyr::rbind.fill(
  metadata[metadata$Vregion  == vregions[i] & 
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))

stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')

```

### PERMANOVA
```{r, fig.width=12, fig.height=6}
permanova <- adonis2(otutab_t ~ DNA_isolation_method, data=sampdat, method=distance)
permanova
permanova.v1v3 <- data.frame(permanova, Vregion=vregions[i])

# Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
print(pairwise_perm)
pairwise_perm.v1v3 <- data.frame(pairwise_perm, Vregion=vregions[i])

```

### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod <- betadisper(dis, groups)
mod
permidisp.v1v3    <- data.frame(distance_to_centroid=mod[["distances"]],  Vregion=vregions[i])
disp.vectors.v1v3 <- data.frame(distance_to_centroid=mod[["vectors"]],
                             Vregion=paste0(vregions[i], ''))

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist.v1v3 <- data.frame(anova(mod), Vregion=vregions[i])
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest.v1v3 <- permutest(mod, pairwise = TRUE, permutations = 999)
permutest.v1v3
permutest.v1v3 <- data.frame(permutest.v1v3[["tab"]], Vregion=vregions[i])
```

### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD <- TukeyHSD(mod))
tukey.disp.v1v3 <- data.frame(mod.HSD[["group"]], Vregion=vregions[i])
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```



## V3-V4
```{r, fig.width=12, fig.height=6}
i <- 3

samples.to.keep <- plyr::rbind.fill(
  metadata[metadata$Vregion  == vregions[i] & 
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))

stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')

```

### PERMANOVA
```{r, fig.width=12, fig.height=6}
permanova <- adonis2(otutab_t ~ DNA_isolation_method, data=sampdat, method=distance)
permanova
permanova.v3v4 <- data.frame(permanova, Vregion=vregions[i])

# Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
print(pairwise_perm)
pairwise_perm.v3v4 <- data.frame(pairwise_perm, Vregion=vregions[i])

```

### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod <- betadisper(dis, groups)
mod
permidisp.v3v4 <- data.frame(distance_to_centroid=mod[["distances"]],  
                             Vregion=vregions[i])
disp.vectors.v3v4 <- data.frame(distance_to_centroid=mod[["vectors"]],
                             Vregion=paste0(vregions[i], ''))

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist.v3v4 <- data.frame(anova(mod), Vregion=vregions[i])
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest.v3v4 <- permutest(mod, pairwise = TRUE, permutations = 999)
permutest.v3v4
permutest.v3v4 <- data.frame(permutest.v3v4[["tab"]], Vregion=vregions[i])
```

### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD <- TukeyHSD(mod))
tukey.disp.v3v4 <- data.frame(mod.HSD[["group"]], Vregion=vregions[i])
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```


## V1-V9 ONT
```{r, fig.width=12, fig.height=6}
i <- 4

samples.to.keep <- plyr::rbind.fill(
  metadata[metadata$platform == 'ONT' & 
           metadata$Vregion  == vregions[i] & 
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))

stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')

```

### PERMANOVA
```{r, fig.width=12, fig.height=6}
permanova <- adonis2(otutab_t ~ DNA_isolation_method, data=sampdat, method=distance)
permanova
permanova.v1v9ont <- data.frame(permanova, 
                                Vregion=paste0(vregions[i], '_ONT'))

# Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
print(pairwise_perm)
pairwise_perm.v1v9ont <- data.frame(pairwise_perm, 
                                    Vregion=paste0(vregions[i], '_ONT'))

```

### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod <- betadisper(dis, groups)
mod
permidisp.v1v9ont <- data.frame(distance_to_centroid=mod[["distances"]], 
                                Vregion=paste0(vregions[i], '_ONT'))
disp.vectors.v1v9ont <- data.frame(distance_to_centroid=mod[["vectors"]],
                               Vregion=paste0(vregions[i], '_ONT'))

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist.v1v9ont <- data.frame(anova(mod), 
                                Vregion=paste0(vregions[i], '_ONT'))
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest.v1v9ont <- permutest(mod, pairwise = TRUE, permutations = 999)
permutest.v1v9ont
permutest.v1v9ont <- data.frame(permutest.v1v9ont[["tab"]],
                                Vregion=paste0(vregions[i], '_ONT'))
```

### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD <- TukeyHSD(mod))
tukey.disp.v1v9ont <- data.frame(mod.HSD[["group"]], 
                                 Vregion=paste0(vregions[i], '_ONT'))
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```


## V1-V9 PacBio
```{r, fig.width=12, fig.height=6}
i <- 4

samples.to.keep <- plyr::rbind.fill(
  metadata[metadata$platform == 'PacBio' & 
           metadata$Vregion  == vregions[i] & 
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))

stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')

```

### PERMANOVA
```{r, fig.width=12, fig.height=6}
permanova <- adonis2(otutab_t ~ DNA_isolation_method, data=sampdat, method=distance)
permanova
permanova.v1v9pb <- data.frame(permanova, 
                               Vregion=paste0(vregions[i], '_PacBio'))

# Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
print(pairwise_perm)
pairwise_perm.v1v9pb <- data.frame(pairwise_perm, 
                                   Vregion=paste0(vregions[i], '_PacBio'))

```

### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod <- betadisper(dis, groups)
mod
permidisp.v1v9pb <- data.frame(distance_to_centroid=mod[["distances"]], 
                               Vregion=paste0(vregions[i], '_PacBio'))

disp.vectors.v1v9pb <- data.frame(distance_to_centroid=mod[["vectors"]],
                               Vregion=paste0(vregions[i], '_PacBio'))

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist.v1v9pb <- data.frame(anova(mod), 
                               Vregion=paste0(vregions[i], '_PacBio'))
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest.v1v9pb <- permutest(mod, pairwise = TRUE, permutations = 999)
permutest.v1v9pb
permutest.v1v9pb <- data.frame(permutest.v1v9pb[["tab"]],
                               Vregion=paste0(vregions[i], '_PacBio'))
```

### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD <- TukeyHSD(mod))
tukey.disp.v1v9pb <- data.frame(mod.HSD[["group"]], 
                                Vregion=paste0(vregions[i], '_PacBio'))
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```


## WGS
```{r, fig.width=12, fig.height=6}
i <- 5

samples.to.keep <- plyr::rbind.fill(
  metadata[metadata$Vregion  == vregions[i] & 
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))

stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')

```

### PERMANOVA
```{r, fig.width=12, fig.height=6}
permanova     <- adonis2(otutab_t ~ DNA_isolation_method, data=sampdat, method=distance)
permanova
permanova.wgs <- data.frame(permanova, Vregion=vregions[i])

# Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
print(pairwise_perm)
pairwise_perm.wgs <- data.frame(pairwise_perm, Vregion=vregions[i])

```

### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis    <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod    <- betadisper(dis, groups)
mod
permidisp.wgs <- data.frame(distance_to_centroid=mod[["distances"]], Vregion=paste0(vregions[i], ''))

disp.vectors.wgs <- data.frame(distance_to_centroid=mod[["vectors"]],
                               Vregion=paste0(vregions[i], ''))

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist.wgs <- data.frame(anova(mod),
                            Vregion=paste0(vregions[i], ''))
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest.wgs <- permutest(mod, pairwise = TRUE, permutations = 999)
permutest.wgs
permutest.wgs <- data.frame(permutest.wgs[["tab"]], 
                            Vregion=paste0(vregions[i], ''))
```

### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD <- TukeyHSD(mod))
tukey.disp.wgs <- data.frame(mod.HSD[["group"]], 
                             Vregion=paste0(vregions[i], ''))
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```


```{r}

permanova <- plyr::rbind.fill(
  data.frame(group=rownames(permanova.v1v2),    permanova.v1v2),
  data.frame(group=rownames(permanova.v1v3),    permanova.v1v3),
  data.frame(group=rownames(permanova.v3v4),    permanova.v3v4),
  data.frame(group=rownames(permanova.v1v9ont), permanova.v1v9ont),
  data.frame(group=rownames(permanova.v1v9pb),  permanova.v1v9pb), 
  data.frame(group=rownames(permanova.wgs),     permanova.wgs) )

pairwise_perm <- plyr::rbind.fill(
  data.frame(group=rownames(pairwise_perm.v1v2),    pairwise_perm.v1v2),
  data.frame(group=rownames(pairwise_perm.v1v3),    pairwise_perm.v1v3),
  data.frame(group=rownames(pairwise_perm.v3v4),    pairwise_perm.v3v4),
  data.frame(group=rownames(pairwise_perm.v1v9ont), pairwise_perm.v1v9ont),
  data.frame(group=rownames(pairwise_perm.v1v9pb),  pairwise_perm.v1v9pb), 
  data.frame(group=rownames(pairwise_perm.wgs),     pairwise_perm.wgs) )

anovadist <- plyr::rbind.fill(
  data.frame(group=rownames(anovadist.v1v2),    anovadist.v1v2),
  data.frame(group=rownames(anovadist.v1v3),    anovadist.v1v3),
  data.frame(group=rownames(anovadist.v3v4),    anovadist.v3v4),
  data.frame(group=rownames(anovadist.v1v9ont), anovadist.v1v9ont),
  data.frame(group=rownames(anovadist.v1v9pb),  anovadist.v1v9pb),
  data.frame(group=rownames(anovadist.wgs),     anovadist.wgs) )

permutest <- plyr::rbind.fill(
  data.frame(group=rownames(permutest.v1v2),    permutest.v1v2),
  data.frame(group=rownames(permutest.v1v3),    permutest.v1v3),
  data.frame(group=rownames(permutest.v3v4),    permutest.v3v4),
  data.frame(group=rownames(permutest.v1v9ont), permutest.v1v9ont),
  data.frame(group=rownames(permutest.v1v9pb),  permutest.v1v9pb), 
  data.frame(group=rownames(permutest.wgs),     permutest.wgs) )

permidisp <- plyr::rbind.fill(
  data.frame(group=rownames(permidisp.v1v2),    permidisp.v1v2),
  data.frame(group=rownames(permidisp.v1v3),    permidisp.v1v3),
  data.frame(group=rownames(permidisp.v3v4),    permidisp.v3v4),
  data.frame(group=rownames(permidisp.v1v9ont), permidisp.v1v9ont),
  data.frame(group=rownames(permidisp.v1v9pb),  permidisp.v1v9pb),
  data.frame(group=rownames(permidisp.wgs),     permidisp.wgs))

tukey.disp <- plyr::rbind.fill(
  data.frame(group=rownames(tukey.disp.v1v2),    tukey.disp.v1v2),
  data.frame(group=rownames(tukey.disp.v1v3),    tukey.disp.v1v3),
  data.frame(group=rownames(tukey.disp.v3v4),    tukey.disp.v3v4),
  data.frame(group=rownames(tukey.disp.v1v9ont), tukey.disp.v1v9ont),
  data.frame(group=rownames(tukey.disp.v1v9pb),  tukey.disp.v1v9pb),
  data.frame(group=rownames(tukey.disp.wgs),     tukey.disp.wgs))

disp.vectors <- plyr::rbind.fill(
  data.frame(group=rownames(disp.vectors.v1v2),    disp.vectors.v1v2),
  data.frame(group=rownames(disp.vectors.v1v3),    disp.vectors.v1v3),
  data.frame(group=rownames(disp.vectors.v3v4),    disp.vectors.v3v4),
  data.frame(group=rownames(disp.vectors.v1v9ont), disp.vectors.v1v9ont),
  data.frame(group=rownames(disp.vectors.v1v9pb),  disp.vectors.v1v9pb),
  data.frame(group=rownames(disp.vectors.wgs),     disp.vectors.wgs))


write_tsv(disp.vectors,   paste0(res.dir, '/disp_vectors.tsv'))
write_tsv(permanova,      paste0(res.dir, '/permanova.tsv'))
write_tsv(pairwise_perm,  paste0(res.dir, '/permanova_pw.tsv'))
write_tsv(permutest,    paste0(res.dir, '/permutest.tsv'))
write_tsv(anovadist,    paste0(res.dir, '/anovadist.tsv'))
write_tsv(permutest,    paste0(res.dir, '/permutest.tsv'))
write_tsv(tukey.disp,   paste0(res.dir, '/tukey.disp.tsv'))
write_tsv(permidisp,    paste0(res.dir, '/permidisp.tsv'))

```



## Combined plots for the results each library separetely

### PCoA
```{r, fig.width=16, fig.height=9}
plot.data <- merge(disp.vectors, metadata, by.x='group', by.y='sample')


cowplot::plot_grid(
  ggplot(plot.data, aes(x=distance_to_centroid.PCoA1, y=distance_to_centroid.PCoA2,
                           color=DNA_isolation_method)) + 
    geom_point(stat="identity") +
    labs(title="",
         y="PCoA_1",
         x="PCoA_2") +
    scale_color_npg() +
    stat_ellipse(aes(group = DNA_isolation_method), type = "norm", level = 0.95) +
    theme_ipsum() +
    theme(panel.spacing.x = unit(2, 'mm'),
          axis.text.x = element_text(angle=90, hjust=1)
          ) +
    facet_grid(cols=vars(Vregion.x)) ,
  
  ggplot(plot.data, aes(x=distance_to_centroid.PCoA1, y=distance_to_centroid.PCoA3,
                           color=DNA_isolation_method)) + 
    geom_point(stat="identity") +
    labs(title="",
         y="PCoA_1",
         x="PCoA_3") +
    scale_color_npg() +
    stat_ellipse(aes(group = DNA_isolation_method), type = "norm", level = 0.95) +
    theme_ipsum() +
    theme(#legend.position = 'none', 
          panel.spacing.x = unit(2, 'mm'),
          axis.text.x = element_text(angle=90, hjust=1)
          ) +
    facet_grid(cols=vars(Vregion.x)) ,
  
  nrow=2)


```

### Boxplot of distance to centroids
```{r, fig.width=12, fig.height=7}
plot.data <- merge(permidisp, metadata, by.x='group', by.y='sample')

ggplot(plot.data, aes(x=DNA_isolation_method, y=distance_to_centroid,
                      color=DNA_isolation_method)) + 
  geom_boxplot() +
  geom_jitter() +
  labs(title="",
       y="",
       x="") +
  scale_color_npg() +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(Vregion.x))

plot.data %>% group_by(DNA_isolation_method, Vregion.x) %>% summarise(mean_distance=mean(distance_to_centroid), sd_distance=sd(distance_to_centroid))

ggplot(plot.data, aes(x=DNA_isolation_method, y=distance_to_centroid,
                      color=DNA_isolation_method)) + 
  geom_boxplot() +
  geom_jitter() +
  labs(title="",
       y="",
       x="") +
  scale_color_npg() +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) 

plot.data %>% group_by(DNA_isolation_method) %>% summarise(mean_distance=mean(distance_to_centroid), sd_distance=sd(distance_to_centroid))
```

### PERMANOVA

PERMANOVA (Permutational Multivariate Analysis of Variance) is a method to test the hypothesis that groups defined by some factor have similar multivariate dispersions and centroids. It's a non-parametric alternative to MANOVA and is particularly suited for community data in ecology. It works by comparing the distance between groups to the distance within groups.

The null hypothesis is that the centroids of the groups in the multivariate space are equal. If the p-value is small (typically below 0.05), we reject the null hypothesis, suggesting the group centroids are indeed different.

```{r}

ggplot(permanova, aes(x=Vregion, y=R2, fill=Vregion)) + 
  geom_bar(stat="identity") +
  labs(title="R-squared values from PERMANOVA across Vregions",
       y="R^2 value",
       x="Vregion") +
  scale_fill_npg() +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(group))


```
The R2 values indicate the proportion of variation explained by the groups. A higher R2 suggests that the grouping factor (in this case, the DNA isolation method) explains a larger proportion of the variation in the data.


```{r}

ggplot(permanova, aes(x=Vregion, y=Pr..F.)) + 
  geom_bar(stat="identity", fill="coral") +
  labs(title="P-values from PERMANOVA across Vregions",
       y="p-value",
       x="Vregion") +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  facet_grid(cols=vars(group))
# Add a line for p=0.05 


```
A p-value less than 0.05, indicate that the centroids of the groups in the multivariate space are significantly different.

### Pair-Wise PERMANOVA


```{r, fig.width=16}

ggplot(pairwise_perm, aes(x=pairs, y=R2, fill=pairs)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=sig)) +
  labs(title="R-squared values from PERMANOVA across Vregions",
       y="R^2 value",
       x="pairs") +
  scale_fill_npg()+
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(Vregion))
```



```{r, fig.width=16}

ggplot(pairwise_perm, aes(x=pairs, y=p.adjusted, fill=pairs)) + 
  geom_bar(stat="identity") +
  geom_text(aes(label=sig)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  labs(title="p-values from pairwise PERMANOVA across Vregions",
       y="R^2 value",
       x="pairs") +
  scale_fill_npg()+
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(Vregion))

```


### PERMUTEST
The permutation test is a non-parametric method used to determine the significance of observed results by comparing them to a distribution of results obtained through random shuffling (or permutations) of the data labels. In the context of community ecology and the PERMANOVA test, a permutation test can be employed to determine if observed differences between groups are significant compared to differences obtained by randomly assigning samples to groups.

The permutation test tested the null hypothesis that the group centroids in the multivariate space are equal. If the p-value is small, it suggests that the observed differences between the groups are significant and not likely due to random chance.
```{r}
ggplot(permutest, aes(x=Vregion, y=F)) + 
  geom_bar(stat="identity", fill="skyblue") +
  labs(title="F-values values from PERMUTEST across Vregions",
       y="F value",
       x="Vregion") +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(group))
```


```{r}
ggplot(permutest, aes(x=Vregion, y=Pr..F.)) + 
  geom_bar(stat="identity", fill="coral") +
  labs(title="P-values values from PERMUTEST across Vregions",
       y="p-value value",
       x="Vregion") +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(group))
```


### ANOVA on and subsequent Tukey post-hoc multiple comparisons of means (95% family-wise confidence level)

```{r}
ggplot(tukey.disp, aes(x=group, y=p.adj, fill=group)) + 
  geom_bar(stat="identity") +
  labs(title="p-values PERMIDISP and subsequent ANOVA and TukeyHSD",
       y="p-value",
       x="comparison") +
  scale_fill_npg() +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(Vregion))


ggplot(tukey.disp) +
  geom_bar(aes(x=group, y=diff, fill=group), stat="identity") +
  geom_pointrange(aes(x=group, y=diff, ymin=lwr, ymax=upr), stat="identity") +
  labs(title="p-values PERMIDISP and subsequent ANOVA and TukeyHSD",
       y="p-value",
       x="comparison") +
  scale_fill_npg() +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(Vregion))

```

### Difference bettween PERMANOVA and Permutation test restuls

```{r}
comparison_df = merge(
  permanova[, c('Vregion', 'F', 'Pr..F.')],
  permutest[, c('Vregion', 'F', 'Pr..F.')],
    by='Vregion', 
    suffixes=c('_PERMANOVA', '_Permutation')
)

# Calculate the absolute differences in F-values and p-values between the two tests
comparison_df['F_difference'] = abs(comparison_df[,'F_PERMANOVA'] - comparison_df[,'F_Permutation'])
comparison_df['p_difference'] = abs(comparison_df[,'Pr..F._PERMANOVA'] - comparison_df[,'Pr..F._Permutation'])

# Display the comparison results
comparison_df[,c('Vregion', 'F_PERMANOVA', 'F_Permutation', 'F_difference', 'Pr..F._PERMANOVA', 'Pr..F._Permutation', 'p_difference')]

```


## Comparison across all DNA isolation methods


### PERMANOVA

Carry out PERMANOVA with combined model (DNA_isolation_kit + Vregion)
```{r, fig.width=12, fig.height=6}
i <- 1

samples.to.keep <- plyr::rbind.fill(
  metadata[#metadata$Vregion  == vregions[i] & 
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

## Add library as another metadata column
ps@sam_data$library <- paste0(ps@sam_data$Vregion, '_', ps@sam_data$platform)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))


stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')
dist_df  <- as.data.frame(as.matrix(ps.dists))

dist_df.gt <- merge(data.frame(sample_1 = rownames(dist_df), dist_df) %>% gather(sample_2, distance, -1),
                    sampdat, by.x='sample_1', by.y='sample')
```

### Full-model PERMANOVA
```{r, fig.width=12, fig.height=6}

permanova.full <- adonis2(otutab_t ~ DNA_isolation_method + library, data=sampdat, method=distance)
permanova.full
permanova.full <- data.frame(group=rownames(permanova.full), permanova.full)  

write_tsv(permanova.full, paste0(res.dir, '/permanova.full.tsv'))
write.table(dist_df,    paste0(res.dir, '/distsance_matrix.tsv'), row.names = T, sep='\t', quote=F)
write.table(dist_df.gt, paste0(res.dir, '/distsance_matrix.gt.tsv'), row.names = T, sep='\t', quote=F)

```

```{r}

ggplot(permanova.full, aes(x=group, y=R2)) + 
  geom_bar(stat="identity", fill="skyblue") +
  labs(title="R-squared values from PERMANOVA across Libraries and DNA isolation kits",
       y="R^2 value",
       x="library") +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(group))
```
The R2 values indicate the proportion of variation explained by the groups. A higher R2 suggests that the grouping factor (in this case, the DNA isolation method) explains a larger proportion of the variation in the data.

```{r}

ggplot(permanova.full, aes(x=group, y=Pr..F.)) + 
  geom_bar(stat="identity", fill="coral") +
  labs(title="P-values from PERMANOVA across Libraries and DNA isolation kits",
       y="p-value",
       x="Vregion") +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  facet_grid(cols=vars(group))# Add a line for p=0.05 


```
A p-value less than 0.05, indicate that the centroids of the groups in the multivariate space are significantly different.


### PairWise PERMANOVA

```{r, fig.width=12, fig.height=6}
# Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
pairwise_perm
pairwise_perm <- data.frame(pairwise_perm)

write_tsv(pairwise_perm, paste0(res.dir, '/pairwise.permanova.tsv'))
```


```{r, fig.width=12, fig.height=9}
ggplot(pairwise_perm, aes(x=pairs, y=R2)) + 
  geom_col(position = "dodge2", fill="skyblue") +
  labs(title="R-squared values from Pairwise PERMANOVA across Library preparation protocols",
       y="R^2 value",
       x="pairs") +
  scale_fill_npg()+
  coord_flip() +
  theme_ipsum() +
  theme(
     #axis.text.x = element_text(angle=45, hjust=1)
    ) 
# + facet_grid(rows=vars(DNA_isolation_method))
```


```{r, fig.width=12, fig.height=9}
ggplot(pairwise_perm, aes(x=pairs, y=p.adjusted)) + 
  geom_col(position = "dodge2", fill="coral") +
  geom_text(aes(label=sig)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  labs(title="p-values from Pairwise PERMANOVA across  Library preparation protocols",
       y="p-value",
       x="pairs") +
  scale_fill_npg()+
  coord_flip() +
  theme_ipsum() +
  theme(
    #axis.text.x = element_text(angle=45, hjust=1)
    )
  # + facet_grid(cols=vars(DNA_isolation_method), scales = 'fixed')

```


### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis    <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod    <- betadisper(dis, groups)
mod

permidisp    <- data.frame(distance_to_centroid=mod[["distances"]])
disp.vectors <- data.frame(distance_to_centroid=mod[["vectors"]])

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist <- data.frame(anova(mod))
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest <- permutest(mod, pairwise = TRUE, permutations = 999)
permutest
permutest.pw <- setNames(data.frame(permutest[["pairwise"]]$permuted), 'permuted_p.value')
permutest.pw <- data.frame(pairs=rownames(permutest.pw), permutest.pw)
permutest <- data.frame(permutest[["tab"]])

```

```{r, fig.width=14, fig.height=9}
ggplot(permutest.pw, aes(x=pairs, y=permuted_p.value)) + 
  geom_bar(stat="identity", fill="coral") +
  labs(title="PERMUTEST: Pair-wise difference in group homogeneity across Libraries",
       y="p-value",
       x="") +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  coord_flip()+
  theme_ipsum() +
  theme(
    #axis.text.x = element_text(angle=0, hjust=0)
    ) #+ facet_grid(cols=vars(group))
```


### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD   <- TukeyHSD(mod))
tukey.disp <- data.frame(mod.HSD[["group"]])
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```



## PERMANOVA on MN and I only

Carry out PERMANOVA with combined model (DNA_isolation_kit + Vregion)
```{r, fig.width=12, fig.height=6}

i <- 1

samples.to.keep <- plyr::rbind.fill(
  metadata[#metadata$Vregion  == vregions[i] & 
           metadata$DNA_isolation_method %in% c('MN', 'I') &
           metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
)$sample

ps <- prune_samples(samples.to.keep, ps.all)

## Add library as another metadata column
ps@sam_data$library <- paste0(ps@sam_data$Vregion, '_', ps@sam_data$platform)

otutab_t <- data.frame(t(otu_table(ps)))
otutab   <- data.frame((otu_table(ps)))
sampdat  <- data.frame(sample_data(ps))


stopifnot(all(rownames(sampdat) == colnames(otutab)))
stopifnot(all(rownames(sampdat) == rownames(otutab_t)))

ps.dists <- phyloseq::distance(ps, distance, type='samples')
dist_df  <- as.data.frame(as.matrix(ps.dists))

dist_df.gt <- merge(data.frame(sample_1 = rownames(dist_df), dist_df) %>% gather(sample_2, distance, -1),
                    sampdat, by.x='sample_1', by.y='sample')

permanova.full <- adonis2(otutab_t ~ DNA_isolation_method + library, data=sampdat, method=distance)
permanova.full
permanova.full <- data.frame(group=rownames(permanova.full), permanova.full)  

## Pairwise PERMANOVA
pairwise_perm <- pairwiseAdonis::pairwise.adonis(ps.dists, sampdat$DNA_isolation_method)
pairwise_perm
pairwise_perm <- data.frame(pairwise_perm)

write_tsv(permanova.full, paste0(res.dir, '/permanova.full.MN.I.tsv'))

```

```{r}

ggplot(permanova.full, aes(x=group, y=R2)) + 
  geom_bar(stat="identity", fill="skyblue") +
  labs(title="R-squared values from PERMANOVA across Libraries and DNA isolation kits",
       y="R^2 value",
       x="library") +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  facet_grid(cols=vars(group))
```
The R2 values indicate the proportion of variation explained by the groups. A higher R2 suggests that the grouping factor (in this case, the DNA isolation method) explains a larger proportion of the variation in the data.


```{r}

ggplot(permanova.full, aes(x=group, y=Pr..F.)) + 
  geom_bar(stat="identity", fill="coral") +
  labs(title="P-values from PERMANOVA across Libraries and DNA isolation kits",
       y="p-value",
       x="Vregion") +
  theme_ipsum() +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  facet_grid(cols=vars(group))# Add a line for p=0.05 


```
A p-value less than 0.05, indicate that the centroids of the groups in the multivariate space are significantly different.


### PERMDISP2
```{r, fig.width=12, fig.height=6}
dis    <- ps.dists
groups <- factor(sampdat$DNA_isolation_method)
mod    <- betadisper(dis, groups)
mod

permidisp    <- data.frame(distance_to_centroid=mod[["distances"]])
disp.vectors <- data.frame(distance_to_centroid=mod[["vectors"]])

```

### ANOVA on test results
```{r, fig.width=12, fig.height=6}
anova(mod)
anovadist <- data.frame(anova(mod))
```

### Permutation test for F
```{r, fig.width=12, fig.height=6}
permutest <- permutest(mod, pairwise = TRUE, permutations = 999)
permutest
permutest.pw <- setNames(data.frame(permutest[["pairwise"]]$permuted), 'permuted_p.value')
permutest.pw <- data.frame(pairs=rownames(permutest.pw), permutest.pw)
permutest <- data.frame(permutest[["tab"]])

```

```{r, fig.width=14, fig.height=9}
ggplot(permutest.pw, aes(x=pairs, y=permuted_p.value)) + 
  geom_bar(stat="identity", fill="coral") +
  labs(title="PERMUTEST: Pair-wise difference in group homogeneity across Libraries",
       y="p-value",
       x="") +
  geom_hline(yintercept=0.05, linetype="dashed", color = "red") +
  coord_flip()+
  theme_ipsum() +
  theme(
    #axis.text.x = element_text(angle=0, hjust=0)
    ) #+ facet_grid(cols=vars(group))
```


### Tukey's Honest Significant Differences
```{r, fig.width=12, fig.height=6}
(mod.HSD   <- TukeyHSD(mod))
tukey.disp <- data.frame(mod.HSD[["group"]])
```

### PCoA and distances to centroid for each group
```{r, fig.width=12, fig.height=6}

par(mfrow = c(1,3))
PCoA1_vs_PCoA2 <- mod

my_cols <- c("#1b9e77", "#7570b3", '#345f92', '#8d2012')

plot(PCoA1_vs_PCoA2, col = my_cols, pch = c(16,17), cex = 1.1, 
     ellipse = TRUE, hull = FALSE)

PCoA1_vs_PCoA3 <- mod
plot(PCoA1_vs_PCoA3, axes = c(3,1), seg.col = "forestgreen", seg.lty = "dashed",
           col = my_cols, pch = c(16,17), cex = 1.1, ellipse = TRUE, hull = FALSE)

### 
#boxplot(mod)
vegan:::boxplot.betadisper(mod)
```



```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```

