---
title: "Toti all samples"
author: Balázs Kakuk
output:
   html_document:
      toc: TRUE
      toc_depth: 3
      toc_float: TRUE
      number_sections: TRUE
      df_print: kable
      
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.align = 'center', fig.height = 10, fig.width = 30, warning = F, message = F, error = F)
library(tidyverse)
library(ggpubr)
library(factoextra)
library(FactoMineR)
library(ggrepel)
library(phyloseq, quietly = T)
library(DESeq2)
library(data.table)
library(ggsci)
library(hrbrthemes)
library(ggh4x)
library(RColorBrewer)
library(stringr); library(stringi)

### Own functions
for(f in list.files('E:/my.R.packages/Rlyeh-main/R', full.names = T)) {source(f)}
for(f in list.files('E:/my.R.packages/minitax/R', full.names = T)) {source(f)}

rownames.from.1st.col <- function(x) {
  x <- data.frame(x[,-1], row.names = x[,1])
  return(x)
}


### General settings

filt.version <- ''
filtering    <- ''
writetables  <- T
save.Figs    <- F
res.dir      <- 'Toti_all_comp';dir.create(res.dir)
#platforms <- c('illumina', 'PacBIO')

## color palette
alpha   <- 0.9
pal.man <- c(brewer.pal(7, 'Set1'), brewer.pal(8, 'Accent')[c(1,2,6)], brewer.pal(8, 'Set2')[c(1,3,5,6,7)], brewer.pal(8, 'Dark2') )
pal.man <- c(colorvec[c(8,32,12:14,17,24,1:2,4,7,5,6,3,10,11,27:28,9,29)], 'grey', 'black')
pal.man <- alpha(c(pal.man, pal.man), alpha)

## taxonomic ranks
ranks <- c("superkingdom", "phylum", "class", "order", "family", "genus", "species")

### minitax settings
## Databse
db <- 'all_NCBI_genomes'
db.dir <- 'E:/data/databases/all_NCBI_genomes'
## Genome Size normalization?
NormToGenomeSize <- F
NormFunction     <- 'mean'
## MAPQ filtering ?
mapq.filt <- NA #  1:59 # 
## Filtering samples
filter.samples.pattern <- 'LCA|Undetermined|SpeciesEstimate'  ### filter out LCA method


## Exclude unclassified reads?
exclude.unclass <- T

## Omit very low abundance taxa
abund.threshold <- 10

### Gold Standard Settings
cami.lineage <- 'db'
add.GS.taxa <- F ## add all GS taxa to the barplots and the correlation plots?
zymo.GS <- read_delim('E:/data/databases/zymo/zymo_6331_composition.txt')
GS_name <- 'GS_reference'
GS_nr   <- 1
just_GS <- T
vars <- 'platform' #('project') # c('DNA_isolation_method', 'db', 'workflow', 'method', 
glom_spec <- T
consider.euk <- T
count_multiplier <- 100000
count_colname    <- 'Genome Copy'

## Fix taxa
crop_taxa <- NULL # 'Veillonella rogosae|Prevotella corporis'

genera_to_glom <- NULL # data.frame(
#  glom= c('Veillonella', 'Faecalibacterium', 'Prevotella'),
#  to  = c('Veillonella rogosae','Faecalibacterium prausnitzii', 'Prevotella corporis')) ## NULL

species_to_glom <- NULL # data.frame(
#  glom= c('Bacteroides fragilis', 'Bacteroides fragilis CAG:47', 'Bacteroides fragilis_A', "Secundilactobacillus pentosiphilus", "Clostridium hathewayi CAG:224"),
#  to  = c(rep('Bacteroides fragilis', 3), 'Lactobacillus pentosiphilus', 'Hungatella hathewayi' )) ## NULL



```



# Notes

- Minitax was run in BestAln method
- In minitax, reads were *NOT* filtered for MAPQ
- In minitax, results were *NOT* normalized to Genome Size
- Unclassified reads *were* excluded
- Taxa with total reads number of 10 or lower were excluded 
- Eukaryotes were *NOT* excluded from the analyis
- Sane samples for different sequencing dates were combined


```{r, import}

#ps.all <- readRDS('Toti_all_PS.summed.filtered.rds')

load(paste0(res.dir, '.RData'))

ps.all <- ps.glom.norm.top.spec

metadata <- data.frame(sample_data(ps.all))
otutab   <- data.frame(otu_table(ps.all))
taxtab   <- data.frame(tax_table(ps.all))

```


# B-diversity Estimations 


Non-metric Multi-Dimensional Scaling


## NMDS on EMU with EMUdb {.tabset}
```{r, echo=FALSE, message=FALSE, warning=FALSE}

####
distance <- "bray" #"robust.aitchison"
norm     <-  NA # 'sum' # NA #


samples.to.keep <- plyr::rbind.fill(rownames(
  metadata[metadata$workflow == 'EMU'   &
           metadata$db       == 'EMUdb', ]
  
))

ps <- prune_samples(samples.to.keep, ps.all)


### 
ps.prop <- norm.ps(ps, norm = norm) ## rlog normalized species counts
## no prior normalization (before distance calculation, there is a normalization)

suppressMessages(ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance=distance))

#nmds.envfit  <- envfit(ord.nmds.bray, sampdat, permutations = 999) # this fits environmental vectors
#nmds.spp.fit <- envfit(ord.nmds.bray, otutab_t, permutations = 999) # this fits species vectors


bray.df <- plot_ordination(ps.prop, ord.nmds.bray, label="sample", title="Bray NMDS", justDF = T)
bray.df$sample_ID <- paste(bray.df$sample_name, bray.df$platform, sep='_')
bray.df$group     <- paste(bray.df$Vregion,     bray.df$platform, sep='_')
```

### Heatmap

```{r, fig.width=16, fig.height=16}
##
ps.sub <- ps

samp_dat <- sample_data(ps.sub)
samp_dat$sample_ID <- paste(samp_dat$platform,
                            samp_dat$Vregion, 
                            samp_dat$sample_name,
                            sep='_')




sample_names(ps.sub) <- samp_dat$sample_ID


ps.dists <- phyloseq::distance(ps.sub, distance, type='samples')
distplot <- fviz_dist(ps.dists, order = F) + theme(axis.text.x = element_text(angle=90, size = 10))

distplot +
  ggtitle('EMU on EMUdb')
#ggsave('Fig4.jpg', distplot, width=12, height=12)


```



### NMDS v1: color = DNA_isolation_method; shape = Vregion
```{r, fig.width=12, fig.height=9}
scatplot <- function(bray.df) {
  ggplot(bray.df, 
         aes(NMDS1, NMDS2)) +
    geom_point(aes(shape      = group, #sequencing_date,
                   color      = DNA_isolation_method), size=2) +
    geom_text_repel(aes(label = sample_name, color=DNA_isolation_method)) +     
    theme_ipsum() 
} 

pn1 <- scatplot(bray.df[,])
pn1
```

### NMDS v2: color = Vregion; shape = DNA_isolation_method
```{r, fig.width=12, fig.height=9}
scatplot <- function(bray.df) {
  ggplot(bray.df, 
         aes(NMDS1, NMDS2)) +
    geom_point(aes(shape      = DNA_isolation_method, #sequencing_date,
                   color      = group), size=2) +
    geom_text_repel(aes(label = sample_name, color=group)) +     
    theme_ipsum() 
} 

pn2 <- scatplot(bray.df[,])
pn2
```




## NMDS on minitax with all_NCBI_genomes {.tabset}


```{r, echo=FALSE, message=FALSE, warning=FALSE}

####
distance <- "bray" #"robust.aitchison"
norm     <-  NA # 'sum' # NA #


samples.to.keep <- plyr::rbind.fill(rownames(
  metadata[metadata$workflow == 'minitax'   &
           metadata$db       == 'all_NCBI_genomes', ]
  
))

ps <- prune_samples(samples.to.keep, ps.all)


### 
ps.prop <- norm.ps(ps, norm = norm) ## rlog normalized species counts
## no prior normalization (before distance calculation, there is a normalization)

suppressMessages(ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance=distance))

#nmds.envfit  <- envfit(ord.nmds.bray, sampdat, permutations = 999) # this fits environmental vectors
#nmds.spp.fit <- envfit(ord.nmds.bray, otutab_t, permutations = 999) # this fits species vectors


bray.df <- plot_ordination(ps.prop, ord.nmds.bray, label="sample", title="Bray NMDS", justDF = T)
bray.df$sample_ID <- paste(bray.df$sample_name, bray.df$platform, sep='_')
bray.df$group     <- paste(bray.df$Vregion,     bray.df$platform, sep='_')
```
### Heatmap

```{r, fig.width=16, fig.height=16}
##
ps.sub <- ps

samp_dat <- sample_data(ps.sub)
samp_dat$sample_ID <- paste(samp_dat$platform,
                            samp_dat$Vregion, 
                            samp_dat$sample_name,
                            sep='_')




sample_names(ps.sub) <- samp_dat$sample_ID


ps.dists <- phyloseq::distance(ps.sub, distance, type='samples')
distplot <- fviz_dist(ps.dists, order = F) + theme(axis.text.x = element_text(angle=90, size = 10))

distplot +
  ggtitle('minitax on all_NCBI_genomes')
#ggsave('Fig4.jpg', distplot, width=12, height=12)


```



### NMDS v1: color = DNA_isolation_method; shape = Vregion
```{r, fig.width=12, fig.height=9}
scatplot <- function(bray.df) {
  ggplot(bray.df, 
         aes(NMDS1, NMDS2)) +
    geom_point(aes(shape      = group, #sequencing_date,
                   color      = DNA_isolation_method), size=2) +
    geom_text_repel(aes(label = sample_name, color=DNA_isolation_method)) +     
    theme_ipsum() 
} 

pn1 <- scatplot(bray.df[,])
pn1
```

### NMDS v2: color = Vregion; shape = DNA_isolation_method
```{r, fig.width=12, fig.height=9}
scatplot <- function(bray.df) {
  ggplot(bray.df, 
         aes(NMDS1, NMDS2)) +
    geom_point(aes(shape      = DNA_isolation_method, #sequencing_date,
                   color      = group), size=2) +
    geom_text_repel(aes(label = sample_name, color=group)) +     
    theme_ipsum() 
} 

pn2 <- scatplot(bray.df[,])
pn2
```

## Invitrogen samples, all DB and software {.tabset}

```{r, echo=FALSE, message=FALSE, warning=FALSE}

####
distance <- "bray" #"robust.aitchison"
norm     <-  NA # 'sum' # NA #


samples.to.keep <- plyr::rbind.fill(rownames(
  metadata[ metadata$DNA_isolation_method == 'I'   
          # metadata$workflow == 'minitax'   
          #&metadata$db       == 'all_NCBI_genomes'
           , ]
  
))

ps <- prune_samples(samples.to.keep, ps.all)


### 
ps.prop <- norm.ps(ps, norm = norm) ## rlog normalized species counts
## no prior normalization (before distance calculation, there is a normalization)

suppressMessages(ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance=distance))

#nmds.envfit  <- envfit(ord.nmds.bray, sampdat, permutations = 999) # this fits environmental vectors
#nmds.spp.fit <- envfit(ord.nmds.bray, otutab_t, permutations = 999) # this fits species vectors


bray.df <- plot_ordination(ps.prop, ord.nmds.bray, label="sample", title="Bray NMDS", justDF = T)
bray.df$sample_ID <- paste(bray.df$sample_name, bray.df$platform, sep='_')
bray.df$group     <- paste(bray.df$Vregion,     bray.df$platform, sep='_')
```

### Heatmap of Bray-Curtis distances

```{r, fig.width=16, fig.height=16}
##
ps.sub <- ps

samp_dat <- sample_data(ps.sub)
samp_dat$sample_ID <- paste(samp_dat$platform,
                            samp_dat$Vregion, 
                            samp_dat$workflow, 
                            samp_dat$db, 
                            samp_dat$sample_name,
                            sep='_')




sample_names(ps.sub) <- samp_dat$sample_ID


ps.dists <- phyloseq::distance(ps.sub, distance, type='samples')
distplot <- fviz_dist(ps.dists, order = F) + theme(axis.text.x = element_text(angle=90, size = 10))

distplot +
  ggtitle('Invitrogen samples, all DB and software')
#ggsave('Fig4.jpg', distplot, width=12, height=12)


```


### NMDS v1: color = DNA_isolation_method; shape = Vregion
```{r, fig.width=12, fig.height=9}
scatplot <- function(bray.df) {
  ggplot(bray.df, 
         aes(NMDS1, NMDS2)) +
    geom_point(aes(shape      = workflow, #sequencing_date,
                   color      = group), size=2) +
    geom_text_repel(aes(label = db, color=group)) +     
    theme_ipsum() 
} 

pn1 <- scatplot(bray.df[,])
pn1
```

### NMDS v2: color = Vregion; shape = DNA_isolation_method
```{r, fig.width=12, fig.height=9}
scatplot <- function(bray.df) {
  ggplot(bray.df, 
         aes(NMDS1, NMDS2)) +
    geom_point(aes(shape      = db, #sequencing_date,
                   color      = group), size=2) +
    geom_text_repel(aes(label = workflow, color=group)) +     
    theme_ipsum() 
} 

pn2 <- scatplot(bray.df[,])
pn2
```


### NMDS v3: color = Vregion; shape = DNA_isolation_method
```{r, fig.width=12, fig.height=9}
scatplot <- function(bray.df) {
  ggplot(bray.df, 
         aes(NMDS1, NMDS2)) +
    geom_point(aes(shape      = group, #sequencing_date,
                   color      = workflow), size=2) +
    geom_text_repel(aes(label = db, color=workflow)) +     
    theme_ipsum() 
} 

pn2 <- scatplot(bray.df[,])
pn2
```


```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```

